---
title: "CHART exercise"
author: "E Poarch"
date: "2025-10-24"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---
**Download packages that need to be used for this script**
```{r}
#install.packages("tableone")
#install.packages("Hmisc")
library(readr)
library(utils)
library(dplyr)
library(knitr)
library(effectsize)
library(tableone)
library(Hmisc)
library(labelled)
```

**Import csv file of data, quick overview to check for missingness and values of each variable**
```{r}
setwd("~/Desktop/CHART_assignment")
dat <- read_csv("data_chart_assignment_2.csv")
describe(dat)
```

**Label variables and specify variable type**
```{r}
unlab_df = dat %>%
  mutate(
    p_code=factor(p_code, labels = c("non-rural","rural")),
    belong_ovr3 = ifelse(community_belong >= 3, 1, 0), #community belonging, scores of 3 or more
    depressed_ovr10 = ifelse(depression_phq8_sum >= 10, 1, 0) #depression score, scores of 10 or more
    )

#label all variables
df = unlab_df %>%
  set_variable_labels(
    p_code="Urban-Rural Postal Code",
    belong_ovr3="Community belonging >=3",
    depressed_ovr10="Depression score >=10",
    depression_phq8_sum="Depression score")
```

**Select variables to be included in district reports for stakeholders**
```{r}
#variables for stratified table by urban/rural
cat_vars <- c("p_code")
cont_vars <- c("depression_phq8_sum")

#variables for overall table including sample size and score cutoffs
overall_vars <- c("belong_ovr3", "depressed_ovr10")
```

**Create folder for reports and sort by unique districts**
```{r}
#create folder for the reports for each district
dir.create("CHART_reports_final", showWarnings = FALSE)
#go through list of unique districts
districts <- unique(df$district)
```

**Create report for each district, includes variables with overall numbers and table stratified by urban/rural**
```{r}
for (district_id in districts) {
  #for loop of each district to create subset and district-specific reports
  district_df <- df %>% filter(district == district_id)
  total_n <- nrow(district_df)
  
  #overall table without stratification
  tabovr <- CreateTableOne(vars = c(overall_vars),
                        data = district_df,
                        factorVars = overall_vars,
                        addOverall = T)
  #extract table as data frame (with variable labels)
  tabovr_df <- as.data.frame(print(tabovr, quote = F, noSpaces = T, printToggle = F,showAllLevels=F,varLabels=T,test=T, smd = F))
  
  #create TableOne stratified by p_code
  tabstrat <- CreateTableOne(vars = c(cat_vars, cont_vars),
                        strata = "p_code",
                        data = district_df,
                        factorVars = cat_vars,
                        addOverall = T)
  #extract table as data frame (with SMD, p-value, variable labels)
  tabstrat_df <- as.data.frame(print(tabstrat, quote = F, noSpaces = T, printToggle = F,showAllLevels=T,varLabels=T,test=T, smd = T))
  
  #create combined table with both overall and stratified
  final_table <- bind_rows(tabovr_df, tabstrat_df, .id = NULL)
  final_df <- as.data.frame(final_table)

  #save as CSV
  write.csv(final_df,
            file = paste0("CHART_reports_final/district_report_", district_id, ".csv"),
            row.names = TRUE)
}
```

**Code for proportion instead of percentage for community belonging and depression (not used)**
```{r}
# prop_belong <- mean(district_df$community_belong >= 3, na.rm = TRUE)
# prop_depressed <- mean(district_df$depression_phq8_sum >= 10, na.rm = TRUE)
#   prop_dat <- data.frame(
#     prop_vbl = c(
#       "prop_belong_ovr3",
#       "prop_depression_ovr10",
#     ),
#     prop_value = c(
#       prop_belong,
#       prop_depressed
#     )
#   )
```

**Zip the reports together**
```{r}
zip("CHART_reports_final.zip", files = list.files("CHART_reports_final", full.names = TRUE))
```
